---
title:
  '01 Analysis - `r unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/"))) - 2]`'
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "UK Dementia Research Institute at UCL"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(
    inputFile,
    encoding = encoding,
    output_file = file.path(
      "..",
      "..",
      "results",
      unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/")))],
      paste0(
        unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/")))], "-01-analysis.html"
      )
    )
  )})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of [Quick Omics](https://github.com/eturkes/quick-omics).*

The data here is derived from @`r unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/")))]` and will be referenced using the name ``r unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/")))]``.

```{r, boilerplate}
# Load in necessary boilerplate and libraries.
# --------------------------------------------

# Copyright 2020, 2022 Emir Turkes, UK DRI at UCL
#
# Licensed under the Apache License, Version 2.0 (the "License");
# you may not use this file except in compliance with the License.
# You may obtain a copy of the License at
#
#     http://www.apache.org/licenses/LICENSE-2.0
#
# Unless required by applicable law or agreed to in writing, software
# distributed under the License is distributed on an "AS IS" BASIS,
# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
# See the License for the specific language governing permissions and
# limitations under the License.

# These should be checked per document.
# -------------------------------------

packages <- c("conflicted", "DESeq2", "DT")
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
source(file.path(getwd(), "..", "utils.R"))

analysis_no <- 1

# Everything else in this chunk remains generally unchanged.
# ----------------------------------------------------------

data_name <- unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/")))] # Name of data.
assets_dir <- normalizePath(file.path(getwd(), "..", "..", "assets", data_name)) # Caches and data.
tmp_dir <- normalizePath(file.path(getwd(), "..", "..", "tmp", data_name))
results_dir <- normalizePath(file.path(getwd(), "..", "..", "results", data_name))

if (!dir.exists(tmp_dir)) {
  dir.create(tmp_dir, recursive = TRUE)
}

# Unique cache directory for each analysis number.
if (!dir.exists(normalizePath(file.path(assets_dir, "cache", paste0("0", analysis_no))))) {
  dir.create(
    normalizePath(file.path(assets_dir, "cache", paste0("0", analysis_no))), recursive = TRUE
  )
}

options(stringsAsFactors = FALSE)
knitr::opts_chunk$set(fig.width = 10, fig.height = 10)
```

# Cell-type/Genotype Interaction Analysis

This analysis looks for interactions between genotype (APOE3 vs. APOE4) and cell-type (neurons vs. astrocytes).
The most relevant tables will be the last two, the first two just looks at cell-type differences in APOE3 and APOE4 alone.
In the first two tables, positive `log2FoldChange` values indicate upregulation in neurons while negative values indicate downregulation, compared to astrocytes.
In the last two tables, positive `log2FoldChange` values indicate upregulation in APOE4 neurons while negative values indicate downregulation, compared to all other combinations of cell-type/genotype groups.
All adjusted p-value cutoffs are set at $0.05$, except the last table.

```{r, prep}
neu <- read.csv(file.path(assets_dir, "GSE102956_COUNTS_NEU.txt"), sep = "\t")
ast <- read.csv(file.path(assets_dir, "GSE102956_COUNTS_AST.txt"), sep = "\t")
data <- cbind(neu, ast)
rm(neu, ast)
coldata <- data.frame(
  batch = factor(rep(seq(0, length(colnames(data)) / 4), each = 3)),
  genotype = factor(rep(c(rep("APOE3", 3), rep("APOE4", 3)), length(colnames(data)) / 6)),
  celltype = factor(c(rep("Neuron", 6), rep("Astrocyte", 6)))
)
dds <- DESeqDataSetFromMatrix(
  data, colData = coldata, design = ~ genotype + celltype + genotype:celltype
)
rm(data)
dds <- DESeq(dds, quiet = TRUE)
```

## Cell-type Comparison in APOE3

```{r, APOE3}
results <- results(
  dds,
  contrast = c(
    unlist(strsplit(resultsNames(dds)[3], "_"))[1],
    unlist(strsplit(resultsNames(dds)[3], "_"))[2],
    unlist(strsplit(resultsNames(dds)[3], "_"))[4]
  )
)
results <- results[which(results$padj < 0.05), ]
results <- results[order(results$padj), ]
datatable_custom(data.frame(results))

n_plots <- 4
for (i in 1:n_plots) {
  barplot(assay(dds)[rownames(results)[i], ], las = 2, main = rownames(results)[i])
}
```

## Cell-type Comparison in APOE4

```{r, APOE4}
results <- results(dds, contrast = list(c(resultsNames(dds)[3], resultsNames(dds)[4])))
results <- results[which(results$padj < 0.05), ]
results <- results[order(results$padj), ]
datatable_custom(data.frame(results))

n_plots <- 4
for (i in 1:n_plots) {
  barplot(assay(dds)[rownames(results)[i], ], las = 2, main = rownames(results)[i])
}
```

## Cell-type/Genotype Interaction

```{r, interaction}
results <- results(dds, name = resultsNames(dds)[4])
results <- results[which(results$padj < 0.05), ]
results <- results[order(results$padj), ]
datatable_custom(data.frame(results))
```

# Neurons Alone

```{r, neuron-prep}
data <- read.csv(file.path(assets_dir, "GSE102956_COUNTS_NEU.txt"), sep = "\t")
coldata <- data.frame(
  batch = factor(rep(seq(0, length(colnames(data)) / 4), each = 3)),
  genotype = factor(rep(c(rep("APOE3", 3), rep("APOE4", 3)), length(colnames(data)) / 6))
)
dds <- DESeqDataSetFromMatrix(data, colData = coldata, design = ~ genotype)
rm(data)
dds <- DESeq(dds, quiet = TRUE)
```

## Upregulation in APOE4

```{r, neuron-APOE4-up}
results <- results(dds, name = resultsNames(dds)[2])
results <- results[which(results$padj < 0.05), ]
results <- results[order(results$padj), ]
datatable_custom(data.frame(results))

n_plots <- 4
for (i in 1:n_plots) {
  barplot(assay(dds)[rownames(results)[i], ], las = 2, main = rownames(results)[i])
}
```

## Upregulation in APOE4 (No Cutoff)

```{r, neuron-APOE4-up-nocutoff}
results <- results(dds, name = resultsNames(dds)[2])
results <- results[order(results$padj), ]
datatable_custom(data.frame(results))
```

# References

This is the concluding section of the document. Here we output the `sessionInfo` and create a bibliography for works cited.

```{r, references}
sessionInfo()
```
