---
title:
  '01 Analysis - `r unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/"))) - 2]`'
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "UK Dementia Research Institute at UCL"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
bibliography:
  '../../`r unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/"))) - 2]`.bib'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(
    inputFile,
    encoding = encoding,
    output_file = file.path(
      "..", "..", "results",
      unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/")))], "01-analysis.html"
    )
  )})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of [Quick Omics](https://github.com/eturkes/quick-omics).*

The data here is derived from @`r unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/")))]` and will be referenced using the name ``r unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/")))]``.

```{r, boilerplate}
# Load in necessary boilerplate and libraries.
# --------------------------------------------

# Copyright 2020 Emir Turkes
#
# Licensed under the Apache License, Version 2.0 (the "License");
# you may not use this file except in compliance with the License.
# You may obtain a copy of the License at
#
#     http://www.apache.org/licenses/LICENSE-2.0
#
# Unless required by applicable law or agreed to in writing, software
# distributed under the License is distributed on an "AS IS" BASIS,
# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
# See the License for the specific language governing permissions and
# limitations under the License.

# These should be checked per document.
# -------------------------------------

packages <- c("conflicted", "DT")
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
source(file.path(getwd(), "..", "utils.R"))

analysis_no <- 1

# Everything else in this chunk remains generally unchanged.
# ----------------------------------------------------------

data_name <- unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/")))] # Name of data.
assets_dir <- normalizePath(file.path(getwd(), "..", "..", "assets", data_name)) # Caches and data.
tmp_dir <- normalizePath(file.path(getwd(), "..", "..", "tmp", data_name))
results_dir <- normalizePath(file.path(getwd(), "..", "..", "results", data_name))

if (!dir.exists(tmp_dir)) {
  dir.create(tmp_dir, recursive = TRUE)
}

# Unique cache directory for each analysis number.
if (!dir.exists(normalizePath(file.path(assets_dir, "cache", paste0("0", analysis_no))))) {
  dir.create(
    normalizePath(file.path(assets_dir, "cache", paste0("0", analysis_no))), recursive = TRUE
  )
}

options(stringsAsFactors = FALSE)
knitr::opts_chunk$set(fig.width = 10, fig.height = 10)
```

# DEGs

```{r, DEG}
data <- read.delim(file.path(assets_dir, "GEO2R.txt"))
data <- data[which(data$adj.P.Val < 0.05), ]
```

## AD Upregulated

```{r, AD-up}
AD_up <- data[which(data$logFC > 0), ]
AD_up <- AD_up[order(-AD_up$logFC), ]
rownames(AD_up) <- NULL
datatable_custom(AD_up)
```

## AD Downregulated

```{r, AD-down}
AD_down <- data[which(data$logFC < 0), ]
AD_down <- AD_down[order(AD_down$logFC), ]
rownames(AD_down) <- NULL
datatable_custom(AD_down)
```

# References

This is the concluding section of the document. Here we output the `sessionInfo` and create a bibliography for works cited.

```{r, references}
sessionInfo()
```
