---
title:
  '01 Analysis - `r unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/"))) - 2]`'
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "UK Dementia Research Institute at UCL"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(
    inputFile,
    encoding = encoding,
    output_file = file.path(
      "..", "..", "results",
      unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/")))], "01-analysis.html"
    )
  )})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of [Quick Omics](https://github.com/eturkes/quick-omics).*

The data here is derived from @`r unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/")))]` and will be referenced using the name ``r unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/")))]``.

```{r, boilerplate}
# Load in necessary boilerplate and libraries.
# --------------------------------------------

# Copyright 2020 Emir Turkes
#
# Licensed under the Apache License, Version 2.0 (the "License");
# you may not use this file except in compliance with the License.
# You may obtain a copy of the License at
#
#     http://www.apache.org/licenses/LICENSE-2.0
#
# Unless required by applicable law or agreed to in writing, software
# distributed under the License is distributed on an "AS IS" BASIS,
# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
# See the License for the specific language governing permissions and
# limitations under the License.

# These should be checked per document.
# -------------------------------------

packages <- c("conflicted", "DESeq2", "DT")
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
source(file.path(getwd(), "..", "utils.R"))

analysis_no <- 1

# Everything else in this chunk remains generally unchanged.
# ----------------------------------------------------------

data_name <- unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/")))] # Name of data.
assets_dir <- normalizePath(file.path(getwd(), "..", "..", "assets", data_name)) # Caches and data.
tmp_dir <- normalizePath(file.path(getwd(), "..", "..", "tmp", data_name))
results_dir <- normalizePath(file.path(getwd(), "..", "..", "results", data_name))

if (!dir.exists(tmp_dir)) {
  dir.create(tmp_dir, recursive = TRUE)
}

# Unique cache directory for each analysis number.
if (!dir.exists(normalizePath(file.path(assets_dir, "cache", paste0("0", analysis_no))))) {
  dir.create(
    normalizePath(file.path(assets_dir, "cache", paste0("0", analysis_no))), recursive = TRUE
  )
}

options(stringsAsFactors = FALSE)
knitr::opts_chunk$set(fig.width = 10, fig.height = 10)
```

# Genotype/Region Interaction Analysis

This analysis looks for interactions between region (entorhinal cortex or EC vs. primary visual cortex or PVC) and genotype (APOE3 vs. APOE4).
The first two tables looks at genotype differences in each region separately, while the last contains the interaction analysis.
In the first two tables, positive `log2FoldChange` values indicate upregulation in APOE4 while negative values indicate downregulation, compared to APOE3.
In the last two tables, positive `log2FoldChange` values indicate upregulation in APOE4 EC while negative values indicate downregulation, compared to all other combinations of region/genotype groups.
All adjusted p-value cutoffs are set at $0.05$, except the last table.

```{r, prep}
ec <- read.csv(file.path(assets_dir, "ec_complete.csv"))
rownames(ec) <- ec[ , 1]
ec <- ec[ , -1]
ec_meta <- read.csv(file.path(assets_dir, "ec_meta.csv"))

pvc <- read.csv(file.path(assets_dir, "pvc_complete.csv"))
pvc <- pvc[ , -1]
pvc_meta <- read.csv(file.path(assets_dir, "pvc_meta.csv"))

data <- cbind(ec, pvc)
rm(ec, pvc)

ec_meta$Sample <- gsub(" ", "", ec_meta$Sample)
ec_meta <- ec_meta[ec_meta$Sample %in% colnames(data), ]
pvc_meta$Sample <- gsub(" ", "", pvc_meta$Sample)
pvc_meta <- pvc_meta[pvc_meta$Sample %in% colnames(data), ]
coldata <- rbind(ec_meta, pvc_meta)
coldata <- data.frame(
  coldata, region = factor(gsub("[[:digit:]]+", "", coldata$Sample)),
  genotype = factor(gsub("/", ".", coldata$Genotype))
)
levels(coldata$region) <- c("PVC", "EC")

dds <- DESeqDataSetFromMatrix(
  data, colData = coldata, design = ~ region + genotype + region:genotype
)
rm(data)
colnames(dds) <- paste0(coldata$region, "-", coldata$genotype)
dds <- DESeq(dds, quiet = TRUE)
```

## APOE4 Upregulation in PVC

```{r, PVC}
genotype_PVC <- results(
  dds,
  contrast = c(
    unlist(strsplit(resultsNames(dds)[3], "_"))[1],
    unlist(strsplit(resultsNames(dds)[3], "_"))[2],
    unlist(strsplit(resultsNames(dds)[3], "_"))[4]
  )
)
res <- genotype_PVC
genotype_PVC <- genotype_PVC[which(genotype_PVC$padj < 0.05), ]
genotype_PVC <- genotype_PVC[order(genotype_PVC$padj), ]
datatable_custom(data.frame(genotype_PVC))

ix <- which.min(res$padj)
res <- res[order(res$padj), ]
barplot(assay(dds)[ix, ], las = 2, main = rownames(dds)[ix])
```

## APOE4 Upregulation in EC

```{r, EC}
genotype_EC <- results(dds, contrast = list(c(resultsNames(dds)[3], resultsNames(dds)[4])))
res <- genotype_EC
genotype_EC <- genotype_EC[which(genotype_EC$padj < 0.05), ]
genotype_EC <- genotype_EC[order(genotype_EC$padj), ]
datatable_custom(data.frame(genotype_EC))

ix <- which.min(res$padj)
res <- res[order(res$padj), ]
barplot(assay(dds)[ix, ], las = 2, main = rownames(dds)[ix])
```

## Region/Genotype Interaction

```{r, interaction}
interaction <- results(dds, name = resultsNames(dds)[4])
interaction <- interaction[which(interaction$padj < 0.05), ]
interaction <- interaction[order(interaction$padj), ]
datatable_custom(data.frame(interaction))
```

## Region/Genotype Interaction (No Cutoff)

This table has no adjusted p-value cutoff applied.

```{r, interaction-nocutoff}
interaction <- results(dds, name = resultsNames(dds)[4])
res <- interaction
interaction <- interaction[order(interaction$padj), ]
datatable_custom(data.frame(interaction))

ix <- which.min(res$padj)
res <- res[order(res$padj), ]
barplot(assay(dds)[ix, ], las = 2, main = rownames(dds)[ix])
```

# References

This is the concluding section of the document. Here we output the `sessionInfo` and create a bibliography for works cited.

```{r, references}
sessionInfo()
```
