---
title:
  '01 Analysis - `r unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/"))) - 2]`'
author:
  - name: "Emir Turkes [emir.turkes@eturkes.com]"
  - name: "UK Dementia Research Institute at UCL"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(
    inputFile,
    encoding = encoding,
    output_file = file.path(
      "..", "..", "results",
      unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/")))], "01-analysis.html"
    )
  )})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
h3 {font-size: 20px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {font-size: 0.95em; padding-left: 25px; text-indent: 0;}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of [Quick Omics](https://github.com/eturkes/quick-omics).*

The data here is derived from @`r unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/")))]` and will be referenced using the name ``r unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/")))]``.

```{r, boilerplate}
# Load in necessary boilerplate and libraries.
# --------------------------------------------

# Copyright 2020 Emir Turkes
#
# Licensed under the Apache License, Version 2.0 (the "License");
# you may not use this file except in compliance with the License.
# You may obtain a copy of the License at
#
#     http://www.apache.org/licenses/LICENSE-2.0
#
# Unless required by applicable law or agreed to in writing, software
# distributed under the License is distributed on an "AS IS" BASIS,
# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
# See the License for the specific language governing permissions and
# limitations under the License.

# These should be checked per document.
# -------------------------------------

packages <- c("conflicted", "maEndToEnd", "biomaRt", "DT")
invisible(suppressPackageStartupMessages(lapply(packages, library, character.only = TRUE)))
source(normalizePath(file.path(getwd(), "..", "utils.R")))
conflict_prefer("strsplit", "Biostrings")
conflict_prefer("rowMedians", "Biobase")
conflict_prefer("summarize", "dplyr")
conflict_prefer("filter", "dplyr")

analysis_no <- 1
color <- colorRampPalette(rev(brewer.pal(n = 7, name = "RdBu")))(100)
color_legend <- c("lightgray", "darkgray")

# Everything else in this chunk remains generally unchanged.
# ----------------------------------------------------------

data_name <- unlist(strsplit(getwd(), "/"))[length(unlist(strsplit(getwd(), "/")))] # Name of data.
assets_dir <- normalizePath(file.path(getwd(), "..", "..", "assets", data_name)) # Caches and data.
tmp_dir <- normalizePath(file.path(getwd(), "..", "..", "tmp", data_name))
results_dir <- normalizePath(file.path(getwd(), "..", "..", "results", data_name))

if (!dir.exists(tmp_dir)) {
  dir.create(tmp_dir, recursive = TRUE)
}

# Unique cache directory for each analysis number.
if (!dir.exists(normalizePath(file.path(assets_dir, "cache", paste0("0", analysis_no))))) {
  dir.create(
    normalizePath(file.path(assets_dir, "cache", paste0("0", analysis_no))), recursive = TRUE
  )
}

options(stringsAsFactors = FALSE)
knitr::opts_chunk$set(fig.width = 10, fig.height = 10)
```

# Prep

First, we run a standard pipeline on the raw data for QC and normalization.

```{r, prep}
sdrf <- read.delim(normalizePath(file.path(tmp_dir, "E-GEOD-4757.sdrf.txt")))
rownames(sdrf) <- sdrf$Array.Data.File
sdrf <- AnnotatedDataFrame(sdrf)

data <- read.celfiles(
  filenames = normalizePath(file.path(tmp_dir, sdrf$Array.Data.File)), phenoData = sdrf,
  verbose = FALSE
)
pData(data) <- pData(data)[, c(
  "Source.Name", "Characteristics.individual.", "Characteristics..clinical.information."
)]

boxplot(data, main = "Boxplot of log2-intensities for Raw Data")

data <- rma(data)

PCA <- prcomp(t(exprs(data)), scale = FALSE)
percentVar <- round(100 * PCA$sdev ^ 2 / sum(PCA$sdev ^ 2), 1)

gg_df <- data.frame(
  PC1 = PCA$x[ , 1],
  PC2 = PCA$x[ , 2],
  Disease = data$Characteristics..clinical.information.,
  Individual = data$Characteristics.individual.
)
gg_df$Individual <- factor(gg_df$Individual)
ggplot(gg_df, aes(PC1, PC2)) +
  geom_point(aes(shape = Individual, colour = Disease)) +
  ggtitle("PCA") +
  xlab(paste0("PC1, VarExp: ", percentVar[1], "%")) +
  ylab(paste0("PC2, VarExp: ", percentVar[2], "%")) +
  scale_shape_manual(values = 1:nlevels(gg_df$Individual)) +
  theme(plot.title = element_text(hjust = 0.5))

boxplot(data, main = "Boxplot of log2-intensities for Normalized Data")

disease_names <- ifelse(
  str_detect(data$Characteristics..clinical.information., "histopathologically"),
  "Normal",
  "NFT"
)
heatmap_anno <- data.frame(Disease = disease_names)
rownames(heatmap_anno) <- rownames(pData(data))
dists <- as.matrix(dist(t(exprs(data)), method = "manhattan"))
rownames(dists) <- rownames(pData(data))
col <- rev(colorRampPalette(brewer.pal(9, "YlOrRd"))(255))
colnames(dists) <- NULL
diag(dists) <- NA
ann_colors <- list(Disease = c(Normal = "blue4", NFT = "cadetblue2"))
pheatmap(
  dists, col = col,
  annotation_row = heatmap_anno, annotation_colors = ann_colors,
  legend = TRUE, treeheight_row = 0,
  legend_breaks = c(min(dists, na.rm = TRUE), max(dists, na.rm = TRUE)),
  legend_labels = (c("small distance", "large distance")), main = "Heatmap for Normalized Data"
)

hist_res <- hist(
  rowMedians(exprs(data)), 100, col = "cornsilk",
  freq = FALSE, main = "Histogram of Median Intensities", border = "antiquewhite4",
  xlab = "Median intensities"
)

print("Remove features not expressed in any cell.")
data <- data[rowSums(exprs(data) > 0) > 0, ]
dim(data)

hist_res <- hist(
  rowMedians(exprs(data)), 100, col = "cornsilk",
  freq = FALSE, main = "Histogram of Median Intensities", border = "antiquewhite4",
  xlab = "Median intensities"
)

exprs(data) <- removeBatchEffect(exprs(data), data$Characteristics.individual.)

annotation(data) <- "hgu133plus2"
data <- nsFilter(data, require.GOBP = TRUE)
data <- data$eset
```

```{r}
rds <- file.path(assets_dir, "cache", paste0("0", analysis_no), "gene_anno.rds")
if (file.exists(rds)) {
  gene_anno <- readRDS(rds)
} else {
  mart <- useEnsembl(
    biomart = "ensembl", dataset = "hsapiens_gene_ensembl", host = "http://useast.ensembl.org/"
  )
  attributes <- c("affy_hg_u133_plus_2", "external_gene_name", "chromosome_name")
  gene_anno <- getBM(
    attributes = attributes, filters = "affy_hg_u133_plus_2",
    values = featureNames(data), mart = mart
  )
  saveRDS(gene_anno, rds)
}

chromosomes <- c(1:22, "MT")
gene_anno <- gene_anno[which(gene_anno$chromosome_name %in% chromosomes), ]
gene_anno <- gene_anno[!duplicated(gene_anno$affy_hg_u133_plus_2), ]
gene_anno <- gene_anno[!duplicated(gene_anno$external_gene_name), ]

data <- data[which(rownames(data) %in% gene_anno$affy_hg_u133_plus_2), ]
fData(data) <- gene_anno
rownames(data) <- fData(data)$external_gene_name

normal <- data[ , grep("histopathologically", data$Characteristics..clinical.information.)]
NFT <- data[ , grep("neurofibrillary", data$Characteristics..clinical.information.)]
```

# DEGs

We calculate DEGs (differentially expressed genes).

```{r, DEG}
# This needs to be edited per comparison.
conditions <- c("Normal", "NFT")
comb_mat <- cbind(exprs(normal), exprs(NFT))
rownames(comb_mat) <- fData(data)$external_gene_name
comb_mat <- comb_mat[!is.na(rownames(comb_mat)), ]
comb_mat <- comb_mat[!duplicated(rownames(comb_mat)), ]
groups <- factor(rep(c(0, 1), c(dim(normal)[2], dim(NFT)[2])))
batch <- c(rep(data$Characteristics.individual.[!duplicated(data$Characteristics.individual.)], 2))

# Set up for `limma` and `heatmap`.
levels(groups) <- conditions
design <- model.matrix(~ groups)
colnames(design) <- conditions
names(color_legend) <- conditions
sample_color_map <- color_legend[groups]

# Fit the linear model and return the top 25 DEGs.
fit <- lmFit(comb_mat, design)
fit <- eBayes(fit)
DE_fit <- topTable(fit, coef = conditions[2], number = Inf, adjust.method = "holm", p.value = 0.05)

# Filter DEGs from the dataset containing our phenotypes of interest.
DE_genes <- comb_mat[rownames(DE_fit), ]

# Make the heatmap.
names(sample_color_map) <- colnames(DE_genes)
gene_set_clustering <- hclust(
  as.dist(1 - cor(t(DE_genes), method = "pearson")), method = "complete"
)
heatmap(
  as.matrix(DE_genes), ColSideColors = sample_color_map,
  xlab = paste0(conditions[2], " vs. ", conditions[1]), margins = c(2, 20),
  col = color, labRow = rownames(DE_genes),
  labCol = " ", Rowv = as.dendrogram(gene_set_clustering),
  Colv = NA
)
legend("topleft", names(color_legend), fill = color_legend, inset = 0.01, bg = "white")

PCA <- prcomp(t(comb_mat), scale = FALSE)
percentVar <- round(100 * PCA$sdev ^ 2 / sum(PCA$sdev ^ 2), 1)

gg_df <- data.frame(PC1 = PCA$x[ , 1], PC2 = PCA$x[ , 2], Disease = groups, Individual = batch)
gg_df$Individual <- factor(gg_df$Individual)
ggplot(gg_df, aes(PC1, PC2)) +
  geom_point(aes(shape = Individual, colour = Disease)) +
  ggtitle("PCA") +
  xlab(paste0("PC1, VarExp: ", percentVar[1], "%")) +
  ylab(paste0("PC2, VarExp: ", percentVar[2], "%")) +
  scale_shape_manual(values = 1:nlevels(gg_df$Individual)) +
  theme(plot.title = element_text(hjust = 0.5))
```

## NFT Upregulated

```{r, NFT-up}
up <- DE_genes[which(DE_fit$logFC > 0), ]
datatable_custom(up)
```

## NFT Downregulated

```{r, NFT-down}
up <- DE_genes[which(DE_fit$logFC < 0), ]
datatable_custom(up)
```

# References

This is the concluding section of the document. Here we output the `sessionInfo` and create a bibliography for works cited.

```{r, references}
sessionInfo()
```
